---
title: "Getting started with egfr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with egfr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(egfr)
```

## Overview

**egfr** provides a vectorised set of equations for estimating glomerular
filtration rate (eGFR) and creatinine clearance from serum creatinine,
cystatin C, or both. This vignette shows the most common workflows.

> These equations are screening tools and are **not** a substitute for measured
> GFR or clinical judgement.

## A single estimate

The CKD-EPI 2021 creatinine equation is the recommended race-free default for
adults:

```{r}
egfr_ckdepi_cr_2021(creatinine = 1.0, age = 50, sex = "female")
```

## Vectorisation

Every argument is vectorised, so you can score a whole cohort in one call:

```{r}
egfr_ckdepi_cr_2021(
  creatinine = c(0.8, 1.2, 1.5),
  age        = c(40, 65, 72),
  sex        = c("female", "male", "female")
)
```

## Working with units

Creatinine defaults to mg/dL. Pass `creatinine_units = "umol/l"` to supply SI
units instead:

```{r}
egfr_ckdepi_cr_2021(88.4, 50, "female", creatinine_units = "umol/l")
```

You can also convert explicitly:

```{r}
convert_creatinine(88.4, from = "umol/l", to = "mg/dl")
```

## Cystatin C and combined equations

If you have cystatin C, or both biomarkers, use the matching function:

```{r}
egfr_ckdepi_cys_2021(cystatin = 1.0, age = 50, sex = "female")
egfr_ckdepi_cr_cys_2021(creatinine = 1.0, cystatin = 1.0, age = 50, sex = "female")
```

## Staging the result

Classify an eGFR value into a KDIGO GFR category:

```{r}
result <- egfr_ckdepi_cr_2021(1.0, 50, "female")
ckd_stage(result)
```

## Body surface area adjustment

Most equations return values normalised to 1.73 m^2. To obtain an absolute
(de-indexed) GFR, compute the body surface area and adjust:

```{r}
area <- bsa(weight = 70, height = 175)
gfr_bsa_adjust(result, bsa = area, to = "absolute")
```

## Where to go next

See the [function reference](https://mhovd.github.io/egfr/reference/) for the
full list of adult, paediatric, and neonatal equations and helpers.
